### 2019-11-08 paper/paper.tex
Work on related works

\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{cvpr}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{multirow} % creates cells that span multiple rows in a table
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage[export]{adjustbox} % adds a frame around figures

% Include other packages here, before hyperref.

% If you comment hyperref and then uncomment it, you should delete
% egpaper.aux before re-running latex.  (Or just hit 'q' on the first latex
% run, let it finish, and you should be clear).
\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=false]{hyperref}

% \cvprfinalcopy % *** Uncomment this line for the final submission

\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}

% Pages are numbered in submission mode, and unnumbered in camera-ready
\ifcvprfinal\pagestyle{empty}\fi
\begin{document}

%%%%%%%%% TITLE
\title{Transform and Tell:\\The News Image Caption Writer That Remembers Rare Names}

\author{Alasdair Tran, Alexander Matthews, Lexing Xie\\
Australian National University\\
{\tt\small \{alasdair.tran,alex.matthews,lexing.xie\}@anu.edu.au}
% For a paper whose authors are all at the same institution,
% omit the following lines up until the closing ``}''.
% Additional authors and addresses can be added with ``\and'',
% just like the second author.
% To save space, use either the email address or home page, not both
}

\maketitle
%\thispagestyle{empty}

%%%%%%%%% ABSTRACT
\begin{abstract}
   We propose an end-to-end model to generate image captions in news articles.
   By combining the transformer architecture, byte-pair encoding, and
   pretrained embeddings from three different modalities (RoBERTa for text,
   ResNet-152 for images, and FaceNet for faces), our system is able to
   describe an image with specific named entities mentioned in the article. Our
   model achieves a CIDEr score of 53 on the GoodNews dataset, significantly
   outperforming the previous state-of-art CIDEr of 13. We also introduce the
   NYTimes+ dataset, the largest news image captioning dataset to date.
   NYTimes+ is an extended version of GoodNews with higher-quality articles and
   metadata that allow us to study the importance of the image location within
   the text. On NYTimes+, we achieve a CIDEr of 51. Pretrained models and
   source code are available from
   \href{https://github.com}{https://github.com/anonymized-link}.
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

Most existing captioning systems can only generate a generic description of an
image \cite{Donahue2015LongTR, Vinyals2015ShowAT, Fang2015FromCT,
Karpathy2015DeepVA, Rennie2017SelfCriticalST, Lu2017KnowingWT,
Anderson2017BottomUpAT, Cornia2019ShowCT}, mainly because these architectures
cannot handle an open vocabulary and early captioning datasets such as MS COCO
\cite{Lin2014MicrosoftCC, Chen2015MicrosoftCC} and Flickr30k
\cite{Young2014FromID} were annotated using only visual cues present in the
image. More recently, datasets such as TIME \cite{Tariq2017ACE}, BreakingNews
\cite{Ramisa2016BreakingNewsAA}, and GoodNews \cite{Biten2019GoodNews} include
real-life captions written by professional journalists. However, existing
models that are trained on these news datasets still need to rely on either
extractive methods or templates to insert named entities.

This motivates us to ask whether it is possible to design a fully end-to-end
model for news image captioning. The model would need to 1) attend to
contextual cues from different modalities and 2) generate captions that contain
specific information including rare proper names that have not been seen in the
training data. To answer this, we look at the recent development in
sequence-to-sequence architectures and language representation learning.

The most successful family of models in the field of natural language
processing (NLP) are the transformer networks, which consist of layers of
self-attention and feed-forward connections stacked on top of each other
\cite{Vaswani2017AttentionIA}. Transformers have consistently been shown to
outperform RNN architectures in language modeling \cite{Radford2019LanguageMA},
story generation \cite{Fan2018HierarchicalNS}, summarization
\cite{Subramanian2019OnEA}, and machine translation \cite{Bojar2018Findings}.
There are theoretical justifications that transformers could be universal
approximations of sequence-to-sequence functions \cite{Anonymous2020AreTU}.

Furthermore, transformers have led to the rise of unsupervised language
representation learning such as BERT \cite{Devlin2019BERT}, XLM
\cite{Lample2019CrosslingualLM}, XLNet \cite{Yang2019XLNetGA}, RoBERTa
\cite{Liu2019RoBERTaAR}, and ALBERT \cite{Lan2019ALBERT}. The contextual word
embeddings provided by these methods have helped establish new state-of-the-art
results on many natural language understanding benchmarks including GLUE
\cite{Wang2019GLUE}, SuperGLUE \cite{Wang2019SuperGLUEAS}, and SQuAD
\cite{Rajpurkar2016SQuAD, Rajpurkar2018KnowWY}, even surpassing the human
baselines in many cases.

\begin{figure}[t]
   \begin{center}
      \includegraphics[width=0.99\linewidth, frame]{figures/figure_1.pdf}
   \end{center}
      \caption{Our transformer model attends over embeddings from three
      different domains (image patches, faces, and article text). Using
      byte-pair encoding, the model can then directly produce a caption
      containing specific named entities without the use of templates.}
   \label{fig:long}
   \label{fig:onecol}
\end{figure}

In parallel to the development of these pre-training methods, many novel
techniques have been proposed to improve training convergence and handle more
diverse data inputs. The most significant contribution is the use of byte-pair
encoding (BPE) \cite{Sennrich2015NeuralMT} to represent a rare word as a
sequence of subword units, thus giving models the ability to handle an open
vocabulary.

The goal of this paper is to make use of these recent developments, along with
their bags of tricks, to carefully design a captioning system most appropriate
for news images. Our main contributions are thus threefold:

\begin{enumerate}
   \item We introduce NYTimes+, the largest news image captioning dataset to
   date, containing 446K articles and 794K images with captions from the New
   York Times spanning 14 years. NYTimes+ is built upon the GoodNews dataset,
   but we write a custom parser to collect higher-quality articles and metadata
   such as the location of an image within the page.

   \item We build a captioning model that combines the power of transformers,
   byte-pair encoding, and attention over three different modalities (text,
   images, and faces). We show that our model achieves state-of-the-art results
   with a significant margin over previous methods, and in particular, it can
   generate names not seen during training without the use of templates.

   \item We provide a detailed model analysis, deconstructing the most
   important modeling components and quantifying the incremental contribution
   that each of them makes not only to the usual metrics such as BLEU, ROUGE,
   METEOR, and CIDEr; but also to other linguistic measures like readability
   scores, caption length, and recall of rare names.
\end{enumerate}

\section{Related Works}

Early image captioning systems mainly used an RNN architecture with a closed
vocabulary \cite{Karpathy2015DeepVA, Donahue2015LongTR, Vinyals2015ShowAT}.
Attention over image patches was later introduced in ``Show, Attend and Tell"
\cite{Xu2015ShowAA}, in which the attention weights are obtained by feeding the
image embeddings and the previous hidden state of the RNN through a multilayer
perception. Later improvements include giving the model the option to not
attend to any image region \cite{Lu2017KnowingWT}, using reinforcement learning
to directly optimize for the CIDEr metric \cite{Rennie2017SelfCriticalST}, and
using a bottom-up approach to propose a region to attend to
\cite{Anderson2017BottomUpAT}. All of these systems can only generate generic
captions with a limited vocabulary.

A more challenging task is to produce captions for news images, in which we
need to be able to generate entity names. Early non-neural approaches include
extractive methods that use n-gram models to combine existing phrases
\cite{Feng2013AutomaticCG} or simply retrieving the most representative
sentence \cite{Tariq2017ACE} in the article. Ramisa \etal
\cite{Ramisa2016BreakingNewsAA} concatenated the word2vec representation of the
article and and the VGG19 representation of the image, and feed them as inputs
to an LSTM generator. However the generator still cannot produce names not
seen in training.

To overcome the limitation of a fixed-size vocabulary, template-based methods
have been used to insert named entities. This involves first generating a
template sentence with placeholders, e.g. ``PERSON speaks at BUILDING in
DATE.'' Afterwards, a selection algorithm is used to pick the best candidate
for each placeholder. Lu \etal \cite{Lu2018EntityAI} built a knowledge graph
for each combination of entities and select the most likely combination.
Meanwhile Biten \etal~\cite{Biten2019GoodNews}, whose GoodNews dataset we will
benchmark against, picked the sentence with the highest cosine similarity with
the template and then found the first entity that matches the type of each
placeholder for insertion. Our proposed model differs from
\cite{Lu2018EntityAI} and \cite{Biten2019GoodNews} in that we are able to
generate a caption with named entities directly without using any intermediate
template.

Transformers are still scarcely used in image captioning. They have been shown
to yield competitive results in generating generic MS COCO captions
\cite{Zhu2018CaptioningTW, Li2019Boosted}. Zhao \etal
\cite{Zhao2019InformativeIC} have gone further and train transformers to
produce named entities in the Conceptual Captions dataset
\cite{Sharma2018ConceptualCA}. However Conceptual Captions have no additional
context apart from the image itself, and the authors use web-entity labels,
extracted using Google Cloud Vision API, as inputs to the model. In our work,
we are more ambitious in that we do not explicitly give the model a list of
entities that should appear in the caption. Instead the model has to determine
which entities to generate on its own by scanning through and attending to the
article.

BPE offers an elegant solution to handle an open vocabulary. To date the only
image captioning work that uses BPE is \cite{Sharma2018ConceptualCA}, but in
their data preprocessing step, they explicitly removed rare named entities from
the captions. Thus no study has been done yet to see how well BPE used in
transformers can generate rare names.

In addition to attending to image patches, some captioning models attend to
objects \cite{Wang2019Hierarchical} and concept terms
\cite{Wang2019Hierarchical,Li2019Boosted}, both of which are derived from the
image itself. When attending to more than one modality, there are various
strategies on how to combine embeddings such as addition, concatenation, and
using multivariate residual modules \cite{Wang2019Hierarchical}.

\section{Datasets}

\subsection{GoodNews and GoodNews+}

As a starting point, we use the GoodNews dataset, previously the largest
dataset on news image captioning \cite{Biten2019GoodNews}. As part of the
quality check, we also reconstruct the same dataset on our own. We use the New
York Times API and write a custom parser to recollect and extract all the
articles and images in the same period as GoodNews from the original source. We
call this GoodNews+. We were able to crawl all the original articles, but the
number of images is slightly less, either due to dead links or not crawling
images unrelated to the main article. GoodNews+ allows us to make the following
observations about the original GoodNews dataset:

\begin{itemize}
   \item The average length of an article in GoodNews is 451 words, while it is
         963 words in GoodNews+. Thus half of the article body is missing in
         GoodNews. This includes the first few paragraphs that might contain
         important information about the top caption. Looking through the
         publicly released code, we found that this is because the original
         parser used by \cite{Biten2019GoodNews} fails to recognize certain
         HTML tags used in New York Times web pages.
   \item GoodNews contains X Spanish articles, adding unwanted noise to the
         training.
   \item GoodNews sometimes include images from the recommendation sidebar,
         which are not part of the main article.
\end{itemize}

Despite the above limitation, we will use the original GoodNews dataset in our
experiments, along with the original train-validation-test split to make
comparison to previous works easier.

\subsection{NYTimes+}

At the same time, we also construct a new dataset, which we call NYTimes+.
NYTimes+ exhibits several advantages over GoodNews:

\begin{itemize}
   \item By increasing the collection period to the last 14 years (March 2005
   -- September 2019), NYTimes+ contains 80\% more articles and images, thus
   becoming the largest news captioning dataset.
   \item NYTimes+ contains only English articles.
   \item We are careful to include only images that are part of the main
   article.
   \item In addition to having a more complete coverage, we also collect
   information about the image position in each article. As shown in Figure
   \ref{fig:dist}, about 40\% of the articles have two or more images.
\end{itemize}

A summary of the datasets is shown in Table \ref{tab:datasets}.

\begin{table}[h]
	\caption {Summary of datasets}
	\label{tab:datasets}
	\centering
	\begin{tabular}{llll}
		\toprule
		  & GoodNews  & GoodNews+ &   NYTimes+ \\
		\midrule
      No. of articles & 241 808 & 241 808 & 445 828 \\
      No. of images   & 462 642 & 440 112 & 794 085 \\
      Article length & 451 & 963 & 974 \\
      Caption length & 18 & 18 & 18 \\
      Start month & Jan 10 & Jan 10 & Mar 05\\
      End month & Jul 18 & Jul 18 & Sep 19 \\
		\bottomrule
	\end{tabular}
\end{table}

\begin{figure}[t]
   \begin{center}
   \includegraphics[width=0.8\linewidth]{figures/caption_dist.png}
   \end{center}
      \caption{Distribution of images across articles.}
   \label{fig:dist}
   \end{figure}

\section{Model}

Devlin \etal \cite{Devlin2019BERT} introduced BERT, a language representation
model which provides pretrained contextual embeddings for text. We use
RoBERTa \cite{Liu2019RoBERTaAR}, a more carefully trained BERT model, to
embed the article text. There there two key advantages:

\begin{itemize}
   \item Unlike glove \cite{Pennington2014Glove} or word2vec
   \cite{Mikolov2013DistributedRO} where each word has exactly one
   representation, the bidirectional and attention mechanism in transformers
   allow words to have different representation depending on the context.
   \item RoBERTa uses Byte-Pair Encoding (BPE)
   \cite{Sennrich2015NeuralMT,Radford2019LanguageMA} which allows us to
   represent and generate potentially any word that can be written in
   Unicode. The vocabulary size is 50,265.
\end{itemize}

Our decoder is a transformer architecture with dynamic convolutions
\cite{Wu2018PayLA} containing four layers with kernel sizes 3, 7, 15 and 31. We
found that using dynamic convolutions allows the training to converge faster
than using the current state-of-the-art GPT-2 architecture
\cite{Radford2019LanguageMA}. We don't have the computational resource to study
the performance of GPT-2.

To make training more efficient, we use adaptive softmax
\cite{Grave2016EfficientSA} and divide the vocabulary into three clusters: 5K,
15K, and 25K. We tie the adaptive weights and we share the decoder input and
output embeddings. We use sinusoidal positional encoding
\cite{Vaswani2017AttentionIA} to represent the position of each token.

We use ResNeXt-101 \cite{Xie2017AggregatedRT} to obtain an embedding for
the images. For each generation step, we use the Scalar-Dot-Product attention
\cite{Vaswani2017AttentionIA} to attend to the image and to article separate.
The two attended representations are then concatenated and fed as input to
the self-attention layer.

We use MTCNN \cite{Zhang2016JointFD} to detect faces and FaceNet
\cite{Schroff2015FaceNetAU} trained on the VGGFace2 dataset
\cite{Cao2017VGGFace2AD}
to classify the faces.

\begin{figure*}[t]
   \begin{center}
   \fbox{\rule{0pt}{2in} \rule{.9\linewidth}{0pt}}
   \end{center}
      \caption{Overall architecture of the model.}
   \label{fig:short}
   \end{figure*}

\section{Experiments}

The training pipeline is written in PyTorch \cite{Paszke2017Automatic} using
the AllenNLP framework \cite{Gardner2017AllenNLP}. The RoBERTa model and
dynamic convolution code comes from fairseq \cite{Ott2019Fairseq}. The code
will made public on GitHub.

Table \ref{tab:results-goodnews} shows the results on the original GoodNews
dataset. Table \ref{tab:results-nytimes} shows additional results done on
our NYTimes+ dataset.

\begin{table*}[t]
	\caption {Results on the GoodNews dataset}
	\label{tab:results-goodnews}
	\centering
	\begin{tabular}{llllllll}
		\toprule
		 & BLEU-1  & BLEU-2 & BLEU-3 & BLEU-4 & ROUGE & METEOR & CIDEr\\
      \midrule
      Biten (Avg + CtxIns) \cite{Biten2019GoodNews} & 9.04 & 3.65 & 1.71 & 0.89 & 12.2 & 4.37 & 13.1 \\
      Biten (Wavg + CtxIns) \cite{Biten2019GoodNews} & 8.10 & 3.33 & 1.60 & 0.85 & 11.9 & 4.23 & 12.8 \\
      Biten (TBB + CtxIns) \cite{Biten2019GoodNews} & 8.44 & 3.42 & 1.62 & 0.87 & 12.2 & 4.30 & 13.1 \\
      Biten (Avg + AttIns) \cite{Biten2019GoodNews} & 8.68 & 3.47 & 1.58 & 0.82 & 12.1 & 4.23 & 12.6 \\
      Biten (Wavg + AttIns) \cite{Biten2019GoodNews} & 7.74 & 3.15 & 1.45 & 0.75 & 11.9 & 4.10 & 12.5 \\
      Biten (TBB + AttIns) \cite{Biten2019GoodNews} & 8.10 & 3.26 & 1.48 & 0.76 & 12.2 & 4.17 & 12.7 \\
      \midrule
      GloVe + LSTM & 13.6 & 6.19 & 3.13 & 1.81 & 13.5 & 5.26 & 11.9 \\
      GloVe + Transformer & 17.2 & 8.68 & 4.81 & 2.91 & 16.4 & 7.13 & 22.2 \\
      RoBERTa + LSTM & 18.7  & 10.3 & 6.09 & 3.89  & 18.0  & 8.19 & 34.8  \\
      RoBERTa + Transformer & 22.2 & 13.4 & 8.68 & 5.99 & 21.2 & 10.1 & 52.9 \\ % Final
		\bottomrule
	\end{tabular}
\end{table*}

\begin{table*}[t]
   \caption {Proper name metrics on the GoodNews test set. We can't compute rare name statistics
   since train-test split were created using a random shuffle.}
	\label{tab:results-names}
	\centering
	\begin{tabular}{lll}
		\toprule
        & \multicolumn{2}{c}{All names}  \\
        & Recall  & Precision \\
      \midrule
      Biten (Avg + CtxIns) \cite{Biten2019GoodNews} & 12.2\% & 16.5\% \\
      Biten (Wavg + CtxIns) \cite{Biten2019GoodNews} & 12.0\% & 17.1\% \\
      Biten (TBB + CtxIns) \cite{Biten2019GoodNews} & 11.7\% & 17.1\% \\
      Biten (Avg + AttIns) \cite{Biten2019GoodNews} & 11.2\% & 17.9\% \\
      Biten (Wavg + AttIns) \cite{Biten2019GoodNews} & 11.2\% & 19.0\% \\
      Biten (TBB + AttIns) \cite{Biten2019GoodNews} & 11.0\% & 19.2\% \\
      \midrule
      GloVe + LSTM & 10.1\% & 14.9\%  \\
      GloVe + Transformer & 16.6\% & 21.7\% \\
      RoBERTa + LSTM & 23.4\% & 28.8\%  \\
      RoBERTa + Transformer & 29.5\% & 33.9\%  \\
		\bottomrule
	\end{tabular}
\end{table*}

\begin{table*}[t]
	\caption {Word statistics on the GoodNews test set}
	\label{tab:results-stats}
	\centering
	\begin{tabular}{lllllll}
		\toprule
        & Caption length  & Unique words \\
      \midrule
      Ground Truths & 18.1 & 17.0 \\
      \midrule
      Biten (Avg + CtxIns) \cite{Biten2019GoodNews}  & 9.9 & 8.8 \\
      Biten (Wavg + CtxIns) \cite{Biten2019GoodNews}  & 9.2 & 8.4 \\
      Biten (TBB + CtxIns) \cite{Biten2019GoodNews}  & 9.2 & 8.4 \\
      Biten (Avg + AttIns) \cite{Biten2019GoodNews}  & 9.7 & 8.4 \\
      Biten (Wavg + AttIns) \cite{Biten2019GoodNews}  & 9.0 & 7.9 \\
      Biten (TBB + AttIns) \cite{Biten2019GoodNews}  & 9.3 & 8.5 \\
      \midrule
      GloVe + LSTM & 13.9 & 11.4 \\
      GloVe + Transformer & 14.9 & 12.9 \\
      RoBERTa + LSTM & 14.7 & 12.7  \\
      RoBERTa + Transformer & 15.5 & 13.7 \\
		\bottomrule
	\end{tabular}
\end{table*}

\begin{table*}[t]
	\caption {Results on the NYTimes+ test set}
	\label{tab:results-nytimes}
	\centering
	\begin{tabular}{llllllll}
		\toprule
		  & BLEU-1  & BLEU-2 & BLEU-3 & BLEU-4 & ROUGE & METEOR & CIDEr \\
      \midrule
      GloVe + LSTM & 14.0 & 6.09 & 2.98 & 1.67 & 13.0 & 5.15 & 9.96 \\
      GloVe + Transformer & 17.0 & 8.42 & 4.63 & 2.79 & 16.1 & 6.99 & 20.6 \\ % Final
      RoBERTa + LSTM & 18.6 & 10.1 & 6.03 & 3.88 & 17.1 & 8.00 & 29.2 \\
      Transformer (first 512 tokens) & 20.7 & 12.4 & 8.14 & 5.73 & 19.8 & 9.54 & 44.1 \\ % Final
      Transformer (location-aware) & 21.7 & 13.3 & 8.84 & 6.25 & 21.3 & 10.3 & 52.4 \\ % Final
      \\
		\bottomrule
	\end{tabular}
\end{table*}

\begin{table*}[t]
	\caption {Proper name metrics on the NYTimes+ test set}
	\label{tab:results-names}
	\centering
	\begin{tabular}{lllllll}
		\toprule
        & \multicolumn{2}{c}{All names} & \multicolumn{2}{c}{Rare names in caption } & \multicolumn{2}{c}{Rare names in article } \\
        & Recall  & Precision & Recall  & Precision & Recall  & Precision \\
      \midrule
      GloVe + LSTM & 10.5\% & 13.1\% & 0\% & 0\% & 0\% & 0\%  \\
      GloVe + Transformer & 18.3\% & 21.7\% & 0\% & 0\% & 0\% & 0\%  \\
      RoBERTa + LSTM & 25.1\% & 26.7\% & 14.8\% & 14.0\% & 12.5\% & 7.98\%  \\
      Transformer (first 512 tokens) & 30.3\% & 32.2\% & 22.6\% & 29.1\% & 20.1\% & 21.3\%  \\
      Transformer (location-aware) & 33.8\% & 36.5\% & 26.2\% & 32.5\% & 24.0\% & 25.6\%   \\
		\bottomrule
	\end{tabular}
\end{table*}

\begin{table*}[t]
	\caption {Word statistics on the NYTimes+ test set}
	\label{tab:results-stats}
	\centering
	\begin{tabular}{lllllll}
		\toprule
        & Caption length  & Unique words \\
      \midrule
      Ground Truths & 18.4 & 17.1 \\
      GloVe + LSTM  & 15.3 & 12.3 \\
      GloVe + Transformer  & 15.1 & 12.9 \\
      RoBERTa + LSTM  & 15.6 & 13.4 \\
      Transformer (first 512 tokens)  & 15.3 & 13.6 \\
      Transformer (location-aware) & 15.1 & 13.5  \\
		\bottomrule
	\end{tabular}
\end{table*}

\begin{table*}[t]
	\caption {Model complexity: Number of trainable parameters}
	\label{tab:models}
	\centering
	\begin{tabular}{lllllll}
		\toprule
        & Parameters \\
      \midrule
      GloVe + LSTM & 157M \\
      GloVe + Transformer & 148M \\
      RoBERTa + LSTM & 159M \\
      RoBERTa + Transformer & 154M \\
        + faces & 171M \\
      %   + names & 175M \\
      %   + faces + names & 192M \\
		\bottomrule
	\end{tabular}
\end{table*}

\section{Conclusion}

{\small
\bibliographystyle{ieee_fullname}
\bibliography{paper}
}

\end{document}