### 2019-11-12 paper/section_0_abstract.tex
Split paper into multiple files

% !TEX root = main.tex

%%%%%%%%% ABSTRACT
\begin{abstract}
    We propose an end-to-end model to generate image captions in news articles.
    By combining the transformer architecture, byte-pair encoding, and
    pretrained embeddings from three different modalities (RoBERTa for text,
    ResNet-152 for images, and FaceNet for faces), our system is able to
    describe an image with specific named entities mentioned in the article. Our
    model achieves a CIDEr score of 54 on the GoodNews dataset, significantly
    outperforming the previous state-of-art CIDEr of 13. We also introduce the
    NYTimes800k dataset, the largest news image captioning dataset to date.
    NYTimes800k is an extended version of GoodNews with higher-quality articles and
    metadata that allow us to study the importance of the image location within
    the text. On NYTimes800k, we achieve a CIDEr of 55. Pretrained models and
    source code are available from
    \href{https://github.com}{https://github.com/anonymized-link}.
 \end{abstract}