### 2019-11-13 paper/section_0_abstract.tex
Cite pointer-generator paper

% !TEX root = main.tex

%%%%%%%%% ABSTRACT
\begin{abstract}
    We propose an end-to-end model to generate image captions in news articles.
    By combining the transformer architecture, byte-pair encoding, copying with
    multi-headed attention, and pretrained embeddings from three different
    modalities (RoBERTa for text, ResNet-152 for images, and FaceNet for
    faces), our system is able to describe an image with specific named
    entities mentioned in the article. Our model achieves a CIDEr score of 54
    on the GoodNews dataset, significantly outperforming the previous
    state-of-art CIDEr of 13. We also introduce the NYTimes800k dataset, the
    largest news image captioning dataset to date. NYTimes800k is an extended
    version of GoodNews with higher-quality articles and metadata that allow us
    to study the importance of the image location within the text. On
    NYTimes800k, we achieve a CIDEr of 55. Pretrained models and source code
    are available from
    \href{https://github.com}{https://github.com/anonymized-link}.
 \end{abstract}